Pipeline systems and infrastructure for the efficient and open processing of Big neuroimaging Data

### 2018-09-27 paper.tex
draft of workflow languages

            To address this issue, some workflow engines have taken the route
            of employing a graphical user interface (GUI). This makes workflow 
            composition accessible to all researchers, including those with
            little to no programming background. The caveat to employing a 
            workflow engine with a GUI interface for workflow composition is
            that users are greatly limited by the tasks they can perform. 
            Workflow engines that employ a worfklow composition GUI include
            SPM, Taverna, LONI and Galaxy. \note{double-check that none have a 
            programming API.Loni does..i think SPM does too}

            As more experienced programmers may feel limited by a GUI, workflow
            engines have decided to provide users with the ability to describe
            their workflows in more flexible formats. Even those with GUIs for 
            workflow composition, such as LONI, enable more experienced users to
            script their workflow. Six different programming
            languages are used for describing scientific workflows: 1) Python, 
            2) MATLAB and Octave, 3) Shell, 4) XML, 6) Common Workflow Language 
            (CWL) and 6) Domain Specific Language (DSL). Workflows engines that 
            employ usage of Python include: Nipype, FastR, Pydpiper, Pegasus
            and Toil.
            Python is an excellent choice as a programming language for 
            scientific workflows, particularly neuroinformatics, due to its 
            widespread use within the community. Other advantages of 
            programming in Python including a wide array of available libraries,
            including neuroimaging-specific ones such as Nilearn and Nibabel
            \note{cite}, and an lower learning curve compared to other 
            programming languages. Additionally, Python is platform independent
            meaning it can be installed on any system. In terms of BigData 
            frameworks that employ Python, Dask is exclusively written in Python
            , whereas Hadoop MapReduce and Spark both have interfaces to Python
            while being written in Java and Scala, respectively.

@@

            Another alternative is to compose workflows using Common Workflow
            Language (CWL). This can be seen in

>>>>

            To address this issue, some workflow engines have taken the route
            of employing a graphical user interface (GUI). This makes workflow 
            composition accessible to all researchers, including those with
            little to no programming background. The caveat to employing a 
            workflow engine with a GUI interface for workflow composition is
            that users are greatly limited by the tasks they can perform. 
            Workflow engines that employ a worfklow composition GUI include
            SPM, Taverna, OpenMOLE, LONI and Galaxy. \note{double-check that 
            none have a programming API.Loni does..i think SPM does too}

            As more experienced programmers may feel limited by a GUI, workflow
            engines have decided to provide users with the ability to describe
            their workflows in more flexible formats. Even those with GUIs for 
            workflow composition, such as LONI and OpenMOLE, allow more 
            experienced users to script their workflow. Six different 
            programming
            languages are used for describing scientific workflows: 1) Python, 
            2) MATLAB and Octave, 3) Shell, 4) XML, 6) Common Workflow Language 
            (CWL) and 6) Domain Specific Language (DSL). Workflows engines that 
            employ usage of Python include: Nipype, FastR, Pydpiper, Pegasus
            and Toil.
            Python is an excellent choice as a programming language for 
            scientific workflows, particularly neuroinformatics, due to its 
            widespread use within the community. Other advantages of 
            programming in Python including a wide array of available libraries,
            including neuroimaging-specific ones such as Nilearn and Nibabel
            \note{cite}, and an lower learning curve compared to other 
            programming languages. Additionally, Python is platform independent
            meaning it can be installed on any system. In terms of BigData 
            frameworks that employ Python, Dask is exclusively written in Python
            , whereas Hadoop MapReduce and Spark both have interfaces to Python
            while being written in Java and Scala, respectively.

@@

            Another alternative is to compose workflows using Common Workflow
            Language (CWL). This can be seen in workflow engines like Toil. 
            CWL is not a programming language, but a specification for workflows
            . Files can either be written in JSON or YAML formats. CWL is 
            based on workflow engines like Make, but differs in that tasks
            are isolated and inputs and outputs must be explicit. Similarly to 
            make, CWL is only available on UNIX platforms. \note{cite cwl 
            website}

            Domain Specific Languages (DSL) are also used to compose workflows.
            DSLs are programming languages that have been adapted to specific
            domains. Workflow engine examples include OpenMOLE, which was 
            derived from Scala, and Netflow, which is derived from the Apache 
            Groovy programming language. While both are written on top of the 
            Java platform, Nextflow is only compatible with UNIX-based 
            environments. These programming languages are not otherwise commonly
            used in neuroimaging, meaning that users must learn specific 
            languages that would not otherwise be used in their field.

            With the exeption of LONI, neuroimaging workflows tend to stick to
            commonly used programming languages such as Python and MATLAB/Octave
            . This is likely due to the fact that the learning curve to adapt to
            these languages is relatively low and users can harness the 
            flexibility of these languages.

            MapReduce, Spark and Dask all provide interfaces to Python, making 
            them good candidates for neuroimaging workflow engines. In addition
            Spark also interfaces with R, another programming language commonly
            used in neuroimaging.
